---
title: "Radial intensity analysis"
author: "Tom van Schaik"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---




### Setup

Load the required libraries.

```{r setup}

# Load libraries
library(pixmap)   ## read.pnm(..)
library(ggplot2)
library(reshape2)
library(ggbeeswarm)

# Prepare output directory
output_dir <- "ts190729_radial_intensity_analysis"
dir.create(output_dir, showWarnings = FALSE)

image_dir <- file.path(output_dir, "images")
dir.create(image_dir, showWarnings = FALSE)

```

```{r knits setup}
library(knitr)
opts_chunk$set(dev=c('png', 'pdf'), fig.path = file.path(output_dir, "figures/"))
pdf.options(useDingbats = FALSE)
```

```{r functions}

ReadPGNFromDF <- function(df, i, target) {
  
  # Get file name
  f <- file.path(image.path,
                 paste0(df$names[i], "_",
                        target,
                        ".pgm"))
  
  # Get image, only grey values
  suppressWarnings(read.pnm(f)@grey*255)
  
}

GetSignalPerLayer <- function(target, dist.factor, f = sum, q = 0.2) {
      
      # Get total signal per layer
      pixellayer <- unlist(tapply(target, dist.factor, f, simplify=FALSE))
      
      # To normalize, subtract background & normalize to 0..1
      d.zero <- quantile(pixellayer, q)
      pixellayer.norm <- pixellayer - d.zero
      pixellayer.norm <- pixellayer.norm / max(pixellayer.norm)
      
      # Area per pixel layer
      pixellayer.area <- table(dist.factor)
      pixellayer.area <- pixellayer.area[pixellayer.area > 0]
      
      list(pixellayer = pixellayer,
           pixellayer.norm = pixellayer.norm,
           pixellayer.area = pixellayer.area)
      
}

TrimMean <- function(x, trim = 0.2) {
  if (length(x) < 3) {
    trim = 0
  }
  y <- x[x >= quantile(x, c(trim)) & x <= quantile(x, c(1 - trim))]
  mean(y)
}

GetHalfwayPoint <- function(d, d.zero, x.max = 40) {
  
  # First, I want to find the lower bound to 0
  # d.zero <- quantile(d, 0.2)
  
  # Then, I want to get rid of the peak itself
  d.subset <- d[1 : (x.max + 5)]
  
  d.max <- which(d.subset > 0.95 * max(d.subset))
  d.max <- as.vector(d.max[length(d.max)])
  #d.subset <- d[(d.max + 1) : length(d)]
  d.subset <- d.subset[(d.max + 1) : length(d.subset)]
  
  # Fit exponential decay function
  z <- nls(y ~ SSasymp(x, yf, y0, log_alpha), 
           data = data.frame(x = seq_along(d.subset), 
                             y = as.vector(d.subset)))
  
  lines(seq_along(d.subset) + d.max, predict(z, seq_along(d.subset)), 
        col = "black", lty = 2)
  
  # Then, what is the halfway value?
  d50 <- mean(c(d.zero, max(d)))
  
  # Can we get this value?
  d50.pix <- approx(x = predict(z), y = seq_along(d.subset) + d.max, xout = d50)$y
  points(d50.pix, d50, col = "black", pch = 19, cex = 2)
  
  # Finally, correct the d50 value for d.zero
  # d50 <- d50 + d.zero
  
  return(list(d50 = d50, d50.pix = d50.pix))
  
}

RadialProfile <- function(data.per.pix, ref.per.pix, area.per.pix, exp.name,
                          dapi.per.pix, lam.per.pix,
                          xlim = c(0, 40), unit = "pixel") {
  
  # profile plot of a signal
  ########################################################################
  if(missing(data.per.pix))
    stop('argument data.per.pix missing')
  if(missing(area.per.pix))
    stop('argument area.per.pix missing')

  # Use shorter names
  d <- data.per.pix
  
  if (missing(dapi.per.pix) == FALSE & missing(lam.per.pix) == FALSE) {
    l <- lam.per.pix
    da <- dapi.per.pix
  }
  
  ### set up plot
  xlab <- paste0("Distance from nuclear rim (", unit, ")")
  ylab <- 'average signal'
  ylim <- c(0, max(c(d)))
  title <- sprintf('profile per %s; %s', unit, exp)
  
  plot(NA, xlim=xlim, ylim=ylim, xlab=xlab, ylab=ylab, 
       ty='o', pch=19, lwd=2, main=title)
  lines(d, col='red', ty='o', pch=19, lwd=2, cex=.5)
  
  if (missing(dapi.per.pix) == FALSE & missing(lam.per.pix) == FALSE) {
    # lines(da, col='blue', ty='o', pch=19, lwd=2, cex=.5)
    lines(l, col='green', ty='o', pch=19, lwd=2, cex=.5)
  } 
  
  # At which pixel do we observe the half-max value, with what value?
  d.zero <- quantile(d, 0.2)
  halfway <- GetHalfwayPoint(d, d.zero = d.zero, x.max = 15)
  d50 <- halfway[["d50"]]
  d50.pix <- halfway[["d50.pix"]]
  
  # Repeat this exercise for the lamina
  if (missing(lam.per.pix) == FALSE) {
    l.zero <- quantile(l, 0.2)
    l.halfway <- GetHalfwayPoint(l, d.zero = l.zero, x.max = 15)
    l.d50 <- l.halfway[["d50"]]
    l.d50.pix <- l.halfway[["d50.pix"]]
  }
  
  # draw lines indicating the d50 values
  lines(x=c(d50.pix, d50.pix), y=c(d50, -1), lty=2, lwd=2, col='grey')
  lines(x=c(l.d50.pix, l.d50.pix), y=c(l.d50, -1), lty=2, lwd=2, col='grey')
  
  # make legend 
  col <- c('red','grey')
  lty <- 1:2
  legend <- c('signal', sprintf('half-max (pixel=%.2f, p50=%.4f)', d50.pix, d50))
  
  if (missing(dapi.per.pix) == FALSE & missing(lam.per.pix) == FALSE) {
    # col <- c('red', 'blue', 'green', 'grey', 'grey')
    # lty <- c(1, 1, 1, 2, 2)
    # legend <- c('tracer', 'dapi', 'lamina', 
    #             sprintf('half-max tracer (pixel=%.2f, p50=%.4f)', d50.pix, d50),
    #             sprintf('half-max lamina (pixel=%.2f, p50=%.4f)', l.d50.pix, l.d50))
    col <- c('red', 'green', 'grey', 'grey')
    lty <- c(1, 1, 2, 2)
    legend <- c('tracer', 'lamina', 
                sprintf('half-max tracer (pixel=%.2f, p50=%.4f)', d50.pix, d50),
                sprintf('half-max lamina (pixel=%.2f, p50=%.4f)', l.d50.pix, l.d50))
  } else {
    col <- c('red','grey')
    lty <- 1:2
    legend <- c('signal', sprintf('half-max (pixel=%.2f, p50=%.4f)', d50.pix, d50))
  }

  legend(x='topr', inset=0.025, bty='n', col=col, lwd=2, lty=lty, legend=legend)
  ret.list <- list(d50 = d50, d50.pix = d50.pix,
                   l.d50 = l.d50, l.d50.pix = l.d50.pix)

  
  # save m.50 values to file
  return(ret.list)
}

```


### Radial profiles

```{r radial profiles}

image.paths <- as.character(c("ts190729_processed_images/ts190221_HTC75_DamID_m6Tracer_analysis/", 
                              "ts190729_processed_images/mv190314_pADam_analysis/",
                              "ts190729_processed_images/ts190411_HTC75_analysis/"))
image.paths <- normalizePath(image.paths)

################################################################################
# MAIN #########################################################################
################################################################################

# loop over the three directories containing images
exp.res <- list()

for (image.path in image.paths) {
  
  # Get image path
  image.path.short <- basename(image.path)
  
  cat('Processing images in', image.path, '\n')
  
  # init results list
  exp.res[[image.path]] <- list()
  
  # List experiments
  experiments <- dir(path=image.path, pattern='segmentation')
  experiments.names <- gsub("_segmentation.pgm", "", experiments)
  
  experiments.df <- data.frame(files = experiments,
                               names = experiments.names,
                               stringsAsFactors = FALSE)
  
  experiments.df$target <- sapply(experiments.df$names, 
                                  function(x) strsplit(x, "_")[[1]][2])
  experiments.df$plate <- sapply(experiments.df$names, 
                                 function(x) strsplit(x, "_")[[1]][3])
  experiments.df$cell <- sapply(experiments.df$names, 
                                function(x) strsplit(x, "_")[[1]][1])
  
  experiments.df$sample <- gsub("_\\d+_\\d+", "", experiments.df$names)
  
  exp.res[[image.path]][["experiments.df"]] <- experiments.df
  
  
  # Loop over experiments
  for (i in 1:nrow(experiments.df)) {
    
    exp <- experiments.df$names[i]
    
    # Read images
    mask <- ReadPGNFromDF(experiments.df, i, "segmentation")
    dist <- ReadPGNFromDF(experiments.df, i, "distancemap")
    dapi <- ReadPGNFromDF(experiments.df, i, "DAPI")
    lamina <- ReadPGNFromDF(experiments.df, i, "lamina")
    tracer <- ReadPGNFromDF(experiments.df, i, "tracer")
    
    # check that distance map is using correct pixel values
    # If not, invert the values
    #######################################################
    # (distance at nuclear lamina = 1)
    tbl <- table(dist)
    if (lm(as.integer(names(tbl))~tbl)[['coefficients']][[2]] > 0) {
      # positive slope; increasing number of pixels with increasing distance values: WRONG!!
      dist <- abs(dist - 255)
    }

    # collect data per pixel layer
    ##############################
    dist.factor <- factor(dist, levels=1:255)
    
    # MEAN meth signal per signal (area = 1 and therefore not used)
    dapi.layer <- GetSignalPerLayer(dapi, dist.factor, f = TrimMean, q = 0.2)
    lamina.layer <- GetSignalPerLayer(lamina, dist.factor, f = TrimMean, q = 0.2)
    tracer.layer <- GetSignalPerLayer(tracer, dist.factor, f = TrimMean, q = 0.2)
    
    # Decide whether the cell passes the quality checks - using the unscaled 
    # data
    t <- tracer.layer[["pixellayer"]]
    
    if (abs(log2(max(t) / quantile(t, 0.2))) < 2) {
      cat(paste0("Skipping sample ", exp, "; no lamina enrichment\n"))
      next
    }
    if (length(t) < 10) {
      cat(paste0("Skipping sample ", exp, "; thickness < 10\n"))
      next
    }
    
    # If all is good, process the image
    cat('Processing image:', exp, '\n')

    # images 
    ########
    # create radial profile plot, and get the d50 distances
    pdf(file.path(image_dir, paste0(image.path.short, "_", exp, "_profile.pdf")),
        width = 7, height = 6)
    
    tryCatch(
    {
      halfway.scores <- RadialProfile(data.per.pix = tracer.layer[['pixellayer.norm']],
                                      area.per.pix = tracer.layer[['pixellayer.area']],
                                      dapi.per.pix = dapi.layer[['pixellayer.norm']],
                                      lam.per.pix  = lamina.layer[['pixellayer.norm']],
                                      exp.name     = exp)
      
      if (! is.na(halfway.scores)[1]) {
        m.50.pix <- halfway.scores[["d50"]]
      } else {
        m.50.pix <- halfway.scores
      }
    },
    error = function(cond) {
      cat("   singular gradient:", exp, "\n")
      halfway.scores <- m.50.pix <- NA
    })
    
    if (any(is.na(halfway.scores))) {
      next
    }
    
    dev.off()
    
    
    # Next, create a image compilation
    png(file.path(image_dir, paste0(image.path.short, "_", exp, "_signal.png")),
        res = 144, width = 900, height = 600)
    opar <- par(mfrow=c(2,3), mar=c(1,1,1,1))
    
    # 1) DAPI
    image(dapi, col=gray(seq(1,0, length.out=256)),axes=FALSE,main="DAPI")
    box()
    
    # 2) Lamina
    image(lamina, col=gray(seq(1,0, length.out=256)), axes=FALSE,main="Lamina")
    box()
    
    # 3) Tracer
    image(tracer, col=gray(seq(1,0, length.out=256)), axes=FALSE,main="Tracer")
    box()
    
    # 4) Mask
    image(mask, col=gray(seq(0,1, length.out=256)), axes=FALSE,main="Mask")
    box()
    
    # 5) Distance map
    image(dist, col=gray(seq(1,0, length.out=256)), axes=FALSE,main="DistanceMap")
    box()
    
    # 6) Halfway distance ( + mask ) 
    image(dist==m.50.pix | dist==1, col=c(0,2),main="m50.pix + dist==1")
    box()
    
    par(opar)
    dev.off()
    
    # Finally, save the values
    exp.res[[image.path]][[exp]] <- list()
    
    exp.res[[image.path]][[exp]][["DAPI"]] <- dapi.layer
    exp.res[[image.path]][[exp]][["lamina"]] <- lamina.layer
    exp.res[[image.path]][[exp]][["tracer"]] <- tracer.layer
    
    exp.res[[image.path]][[exp]][["halfway.scores"]] <- halfway.scores
    
  }
}

```


### Combine results per experiment

The previous section simply reads all the files and processes them in the same
way. This means that individual experiments are not treated differently. Here,
I will generate plots for this experiment specifically. 

```{r combine results per experiment, fig.width = 6, fig.height = 5}

# Prepare output
classes.list <- list()

for (image.path in image.paths) {
  
  # Get combined experiment.df
  experiments.df <- exp.res[[image.path]][["experiments.df"]]
  
  # Get classes
  filter <- c("Damonly", "neg", "tryp", "withoutshield", "shield_pos")
  
  # classes <- unique(paste(experiments.df$cell,
  #                         experiments.df$plate,
  #                         experiments.df$target,
  #                         sep = "_"))
  classes <- unique(experiments.df$sample)
  classes <- grep(paste(filter, collapse = "|"), classes, invert = T, value = T)
  
  # Define range of interest
  xlim <- c(0, 30)
  df.template <- data.frame(distance = 1:xlim[2])
  
  for (class in classes) {
    
    # Get the samples for every experiment
    samples <- grep(class, experiments.df$names, value = T)
    samples.filter <- c()
    
    # Get signal for every pixel
    df <- data.frame(distance = c(1:xlim[2]))
    
    for (exp in samples) {
      
      if (is.null(exp.res[[image.path]][[exp]])) {
        next
      }
      samples.filter <- c(samples.filter, exp)
      
      m <- exp.res[[image.path]][[exp]][["tracer"]][["pixellayer.norm"]]
      d <- exp.res[[image.path]][[exp]][["DAPI"]][["pixellayer.norm"]]
      l <- exp.res[[image.path]][[exp]][["lamina"]][["pixellayer.norm"]]
      
      df.tmp <- data.frame(distance = names(m), m = m, l = l, d = d)
      df <- merge(df, df.tmp, by = "distance", all = T)
      
      names(df)[(ncol(df) - 2) : ncol(df)] <- paste0(exp, "_",
                                                     c("tracer", "lam", "dapi"))
    
    }
    
    df$distance <- as.numeric(df$distance)
    df <- df[order(df$distance), ]
    
    # Calculate mean signal for every pixel
    df$tracer <- rowMeans(df[, paste0(samples.filter, "_tracer")])
    df$lam <- rowMeans(df[, paste0(samples.filter, "_lam")])
    df$dapi <- rowMeans(df[, paste0(samples.filter, "_dapi")])
    
  
    # Plotting
    xlab <- 'pixels away from nuclear rim'
    ylab <- 'average signal per pixel'
    
    plot(NA, xlim=xlim, ylim=c(0, 1), xlab=xlab, ylab=ylab, ty='o', pch=19, lwd=2,
         main = class)
    lines(df$tracer, col='red', ty='o', pch=19, lwd=2, cex=.5)
    # lines(dapi, col='blue', ty='o', pch=19, lwd=2, cex=.5)
    lines(df$lam, col='green', ty='o', pch=19, lwd=2, cex=.5)
    # lines(x=c(-10, m.50.pix, m.50.pix), y=c(m.50, m.50, -1), lty=2, lwd=2, col='grey')
    # legend(x='topr', inset=0.025, bty='n', col=c('red', 'green', 'grey'), lwd=2, lty=c(1,1,2),
    #        legend=c('methyl-tracer', 'lamina', sprintf('half-max (pixel=%i, p50=%.4f)', m.50.pix, m.50)))
    legend(x='topr', inset=0.025, bty='n', col=c('red', 'green', 'grey'), lwd=2, lty=c(1,1,2),
         legend=c('methyl-tracer', 'lamina'))
    
    classes.list[[class]][["df"]] <- df
  
  }
}





# Combine lines in a single plot
classes <- names(classes.list)

# 1) Antibody
plot(NA, xlim=xlim, ylim=c(0, 1), xlab=xlab, ylab=ylab, ty='o', pch=19, lwd=2,
     main = "Lamina distributions")

for (i in seq_along(classes)) {

  class <- classes[i]
  lines(classes.list[[class]][["df"]][, "lam"], col=i, ty='o', pch=19, lwd=2, cex=.5)
}

legend(x='topright', inset=0.025, bty='n', col = 1:length(classes), lwd=2,
       legend=classes)


# 2) Tracer
plot(NA, xlim=xlim, ylim=c(0, 1), xlab=xlab, ylab=ylab, ty='o', pch=19, lwd=2,
     main = "Tracer distributions")

for (i in seq_along(classes)) {

  class <- classes[i]
  lines(classes.list[[class]][["df"]][, "tracer"], col=i, ty='o', pch=19, lwd=2, cex=.5)
}

legend(x='topright', inset=0.025, bty='n', col = 1:length(classes), lwd=2,
       legend=classes)


# 3) DAPI
plot(NA, xlim=xlim, ylim=c(0, 1), xlab=xlab, ylab=ylab, ty='o', pch=19, lwd=2,
     main = "DAPI distributions")

for (i in seq_along(classes)) {

  class <- classes[i]
  lines(classes.list[[class]][["df"]][, "dapi"], col=i, ty='o', pch=19, lwd=2, cex=.5)
}

legend(x='topright', inset=0.025, bty='n', col = 1:length(classes), lwd=2,
       legend=classes)


# Finally, combined antibody + tracer in one figure
# Combine lines in a single plot
plot(NA, xlim=xlim, ylim=c(0, 1), xlab=xlab, ylab=ylab, ty='o', pch=19, lwd=2,
     main = "Lamina & tracer distributions")

for (i in seq_along(classes)) {

  class <- classes[i]
  lines(classes.list[[class]][["df"]][, "lam"], col=i, ty='o', pch=19, lwd=2, cex=.5, lty = 2)
  lines(classes.list[[class]][["df"]][, "tracer"], col=i, ty='o', pch=19, lwd=2, cex=.5)
}

legend(x='topright', inset=0.025, bty='n', col = 1:length(classes), lwd=2,
       legend=classes)
legend(x='bottomleft', inset=0.025, bty='n', col = 1, lwd=2, lty = c(2, 1),
       legend=c("lamina", "tracer"))

```


### Per-cell difference in m50 point

Alternatively, I can look at the difference in m50 point for the lamina and 
tracer, for every cell. And put these into a scatterplot.

```{r m50 differences, fig.width = 4.5, fig.height = 5.5}

d50 <- c()
classes.selection <- classes

for (class in classes.selection) {
  
  # Combine experiments.df
  for (image.path in image.paths) {
    
    # Get combined experiment.df
    experiments.df <- exp.res[[image.path]][["experiments.df"]]
    
    # Get the samples for every experiment
    samples <- grep(class, experiments.df$names, value = T)
    
    # Get a vector of halfway points
    for (exp in samples) {
      
      if (is.null(exp.res[[image.path]][[exp]])) {
        next
      }
      
      d50.sample <- exp.res[[image.path]][[exp]][["halfway.scores"]][["d50.pix"]]
      l50.sample <- exp.res[[image.path]][[exp]][["halfway.scores"]][["l.d50.pix"]]
      
      d50 <- rbind(d50, 
                   data.frame(class = class, exp = exp, 
                              d50 = d50.sample, l50 = l50.sample,
                              diff = d50.sample - l50.sample,
                              image.path = basename(image.path)))
      
    }
  }
}

# Manually: filter classes further
d50$class[d50$class == "HT1080_LMNB2"] <- "HT1080_pADamID_LMNB2"
d50 <- d50[d50$class != "RPE_LMNB2", ]

plt <- ggplot(d50, aes(x = class, y = diff)) +
  geom_boxplot(outlier.shape = NA, fill = NA, width = 0.2) +
  geom_quasirandom() +
  xlab("") +
  ylab("Difference (pixel)") +
  ggtitle("Difference halfway decay tracer vs lamina") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

plot(plt)

plt <- ggplot(d50, aes(x = class, y = diff, col = image.path)) +
  geom_boxplot(outlier.shape = NA, fill = NA, width = 0.2) +
  geom_quasirandom() +
  xlab("") +
  ylab("Difference (pixel)") +
  ggtitle("Difference halfway decay tracer vs lamina") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

plot(plt)

```


### Conclusion

First of all, this needs a bit more work and a few more cells to be really 
reliable. Still, we can have some observations:

  * The trypsin sample (very ugly cells) and the -shield sample (very little
    signal) are probably useless and can be disregarded for now.
  * There is almost no difference between the flat cells and the nuclear 
    isolated cells with regard to their lamina and tracer "thickness". 
  * For the pA-DamID sample, we can see that the lamina and the tracer are 
    even less thick. Especially for the lamina, I don't know why. Most likely,
    we have very few cells and were a little bit more stringent in selecting 
    useful cells. Also, the cells have less of these invaginations and lamina
    bridges. Coindicide or pA-DamID?
  * Regardless of the thinner lamina, we can definitely say that the tracer is 
    not thicker. If anything, it is even thinner than the shield induced sample.
    (Which actually makes sense, right? There are no active movements in pA-
    DamID.)
    
I think we should repeat this exercise, preferably a bit cleaner, and leave it
at this.


### SessionInfo

```{r sessionInfo}

sessionInfo()

```


