---
title: "pA-DamID correlate enrichment"
author: "Tom van Schaik"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---

I have determined LAD enrichment for several pA-DamID experiments using both
microscopy and sequencing. The question now: how well do these values correlate?

### Setup

Load the required libraries.

```{r setup}

# Load libraries
library(ggplot2)
library(reshape2)
library(ggbeeswarm)
library(RColorBrewer)

```

### Define functions

Functions required for this document.

```{r define functions}

output_dir <- "ts190727_pADamID_correlate_enrichments"

```

```{r knits setup}
library(knitr)
opts_chunk$set(fig.width = 10, fig.height = 4, 
               dev=c('png', 'pdf'), fig.path = file.path(output_dir, "figures/"))
pdf.options(useDingbats = FALSE)
```

### Load data

Read in the data.

```{r load data}

# Input directory for the files
enrichment_microscope <- read.table("ts190729_all_enrichments_microscope.txt",
                                    sep = "\t", header = T)
enrichment_sequencing <- read.table("ts190729_LADenrichment/lad_enrichment.txt",
                                    sep = "\t", header = T)

```

### Link data sets

Unfortunately, the names do not really match between the experiments. I need
to manually link them.

```{r link data sets}

enrichment_microscope$sequencing_name <- c("pADamID-Hap1_r2_LMNB1-rabbit",
                                           "pADamID-Hap1_r2_LMNB2", 
                                           "pADamID-Hap1_r2_LMNB2-no2ndrabbit",
                                           NA,
                                           "pADamID-Hap1_r2_NUP153",
                                           "pADamID-Hap1_r2_pADam",
                                           "pADamID-Hap1_r2_2ndrabbit-pADam",
                                           NA,
                                           "pADamID-Hap1_r1_LMNB1-BSA",
                                           NA,
                                           "pADamID-Hap1_r1_H3K27me3",
                                           "pADamID-Hap1_r1_H3K9me2",
                                           "pADamID-Hap1_r1_H3K9me3",
                                           "pADamID-Hap1_r1_LMNB1-500d",
                                           "pADamID-Hap1_r1_LMNB1-500d-pADam-500d",
                                           "pADamID-Hap1_r1_Dam",
                                           "pADamID-Hap1_r1_pADam",
                                           NA,
                                           "pADamID-Hap1_r1_LMNB1",
                                           NA,
                                           NA,
                                           NA,
                                           "pADamID-Hap1_r1_LMNB1-RT",
                                           "pADamID-Hap1_r1_LMNB1-10m",
                                           NA,
                                           "pADamID-Hap1_r4_Dam_semibad-Dam",
                                           "pADamID-Hap1_r4_H3K27me3",
                                           "pADamID-Hap1_r4_H3K9me3",
                                           "pADamID-Hap1_r4_LMNB1-100",
                                           "pADamID-Hap1_r4_LMNB1-500",
                                           "pADamID-Hap1_r4_LMNB2-100",
                                           "pADamID-Hap1_r4_LMNB2-500",
                                           "pADamID-Hap1_r4_NUP153",
                                           "pADamID-Hap1_r4_pADam")

enrichment_combined <- merge(enrichment_microscope, enrichment_sequencing,
                             by.x = "sequencing_name", by.y = "samples", all = FALSE)

# Subset different types of experiments
enrichment_combined$group <- NA
enrichment_combined$group[grep("_Dam", enrichment_combined$sequencing_name)] <- "Dam"
enrichment_combined$group[grep("_pADam|2ndrabbit-pADam", enrichment_combined$sequencing_name)] <- "no-primary"
enrichment_combined$group[grep("LMNB1", enrichment_combined$sequencing_name)] <- "Lamin B1"
enrichment_combined$group[grep("LMNB2", enrichment_combined$sequencing_name)] <- "Lamin B2"
#enrichment_combined$group[grep("NUP", enrichment_combined$sequencing_name)] <- "nuclear-pore"
enrichment_combined$group[grep("H3K27", enrichment_combined$sequencing_name)] <- "H3K27me3"
enrichment_combined$group[grep("H3K9me3", enrichment_combined$sequencing_name)] <- "H3K9me3"
enrichment_combined$group[grep("LMNB2-no2ndrabbit", enrichment_combined$sequencing_name)] <- "no-bridging-antibody"
enrichment_combined$group <- factor(enrichment_combined$group,
                                    levels = c("Dam", "Lamin B1", "Lamin B2", "nuclear-pore", 
                                               "H3K27me3", "H3K9me3", "no-primary",
                                               "no-bridging-antibody"))

```

### Plot enrichment

Finally, create a plot with enrichment scores. 

```{r plot enrichment, fig.width = 5, fig.height = 3.5}

enrichment_combined <- enrichment_combined[order(enrichment_combined$ratio, decreasing = T), ]
enrichment_combined

pearson <- round(cor(enrichment_combined$enrichment, enrichment_combined$ratio),
                 digits = 2)
limits <- range(c(enrichment_combined$enrichment, enrichment_combined$ratio))

plt1 <- ggplot(enrichment_combined, aes(x = enrichment, y = ratio)) +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", col = "red") +
  ggtitle(paste("Correlate microscopy and sequencing - Pearson:", pearson)) +
  xlab("Enrichment microscope") +
  ylab("Enrichment sequencing") +
  xlim(limits[1], limits[2]) +
  ylim(limits[1], limits[2]) +
  theme_bw() +
  coord_fixed()

plt2 <- ggplot(enrichment_combined, aes(x = enrichment, y = ratio, label = sequencing_name)) +
  geom_point() +
  geom_text(aes(y = ratio + 0.1)) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", col = "red") +
  ggtitle(paste("Correlate microscopy and sequencing - Pearson:", pearson)) +
  xlab("Enrichment microscope") +
  ylab("Enrichment sequencing") +
  xlim(limits[1], limits[2]) +
  ylim(limits[1], limits[2]) +
  theme_bw() +
  coord_fixed()

plt3 <- ggplot(enrichment_combined, aes(x = enrichment, y = ratio)) +
  geom_point(aes(col = group)) +
  geom_smooth(method = "lm") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", col = "red") +
  ggtitle(paste("Correlate microscopy and sequencing - Pearson:", pearson)) +
  xlab("Enrichment microscope") +
  ylab("Enrichment sequencing") +
  xlim(limits[1], limits[2]) +
  ylim(limits[1], limits[2]) +
  scale_color_brewer(palette = "Set1") +
  theme_bw() +
  coord_fixed()

plot(plt1)
plot(plt2)
plot(plt3)

pdf("ts181127_pADamID_correlate_enrichments.pdf", width = 6, height = 4)
plot(plt3)
dev.off()

```

### SessionInfo

```{r sessionInfo}

sessionInfo()

```


