---
title: "pA-DamID peripheral tracer enrichment"
author: "Tom van Schaik"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---

pA-DamID peripheral tracer enrichment.

Given 4 .csv files for every slide:

  * periphery, tracer
  * interior, tracer
  * periphery, antibody
  * interior, antibody
  
determine the peripheral enrichment. Note that these files contain
coordinates that should be matched - the order can differ between them!

### Setup

Load the required libraries.

```{r setup}

# Load libraries
library(ggplot2)
library(reshape2)
library(ggbeeswarm)
library(RColorBrewer)
library(tidyverse)

# Output file
output_file <- "ts200526_all_enrichments.txt"
output_dir <- "ts200526_pADamID_peripheral_tracer_enrichment_combined"
dir.create(output_dir, showWarnings = FALSE)

```

```{r knits setup}
library(knitr)
opts_chunk$set(dev=c('png', 'pdf'), fig.path = file.path(output_dir, "figures/"))
pdf.options(useDingbats = FALSE)
```


### Define functions

Functions required for this document.

```{r define functions}

ListSlides <- function(directory) {
  # Given an input directory, list all the slide names in here.
  files <- list.files(directory)
  files <- grep("periphery,tracer", files, value = T)
  
  slides <- gsub(",periphery,tracer.csv", "", files)
  slides
}

GetExperiment <- function(slides) {
  # Given a vector of slide names, return a factor of experiment names
  experiments <- factor(gsub("_[0-9]+$", "", slides))
  experiments
}

ReadSlide <- function(slide_name, experiment, max_distance = 10) {
  # Given a slide name, return a data frame with mean intensities. Combine the 
  # various .csv files using their coordinates.
  
  # Note: with the HALO Tracer, these are in a different order
  periphery_tracer <- read.csv(paste0(slide_name, ",periphery,tracer.csv"))
  interior_tracer <- read.csv(paste0(slide_name, ",interior,tracer.csv"))
  periphery_antibody <- read.csv(paste0(slide_name, ",periphery,antibody.csv"))
  interior_antibody <- read.csv(paste0(slide_name, ",interior,antibody.csv"))
  # periphery_tracer <- read.csv(paste0(slide_name, ",periphery,antibody.csv"))
  # interior_tracer <- read.csv(paste0(slide_name, ",interior,antibody.csv"))
  # periphery_antibody <- read.csv(paste0(slide_name, ",periphery,tracer.csv"))
  # interior_antibody <- read.csv(paste0(slide_name, ",interior,tracer.csv"))
  
  # For every cell, find the matching cells in the other data frames
  slide <- c()
  for (i in 1:nrow(periphery_tracer)) {
    periphery_tracer_row <- periphery_tracer[i, ]
    
    periphery_tracer_value <- periphery_tracer_row["Mean"]
    interior_tracer_value <- interior_tracer[which(abs(interior_tracer$X - periphery_tracer_row$X) < max_distance &
                                                   abs(interior_tracer$Y - periphery_tracer_row$Y) < max_distance),
                                             "Mean"]
    periphery_antibody_value <- periphery_antibody[which(abs(periphery_antibody$X - periphery_tracer_row$X) < max_distance &
                                                         abs(periphery_antibody$Y - periphery_tracer_row$Y) < max_distance),
                                                   "Mean"]
    interior_antibody_value <- interior_antibody[which(abs(interior_antibody$X - periphery_tracer_row$X) < max_distance &
                                                       abs(interior_antibody$Y - periphery_tracer_row$Y) < max_distance),
                                                 "Mean"]
    surface <- periphery_tracer_row[, 2] + 
      interior_tracer[which(abs(interior_tracer$X - periphery_tracer_row$X) < max_distance &
                              abs(interior_tracer$Y - periphery_tracer_row$Y) < max_distance),
                      2]
    
    if (length(periphery_tracer_value) == 1 &
          length(interior_tracer_value) == 1 &
          length(periphery_antibody_value) == 1 &
          length(interior_antibody_value) == 1 &
          length(surface) == 1) {
     
      slide <- rbind(slide,
                     data.frame(periphery_tracer_value,
                                interior_tracer_value,
                                periphery_antibody_value,
                                interior_antibody_value,
                                surface = surface,
                                experiment,
                                stringsAsFactors = FALSE)) 
      
    }

  }
  
  # Make sure the names are correct
  names(slide) <- c("periphery_tracer",
                    "interior_tracer",
                    "periphery_antibody",
                    "interior_antibody",
                    "surface",
                    "experiment")
  
  slide
}

CalculateLog2Ratios <- function(data) {
  # Calculate the enrichment ratios (periphery over interior) for all the rows
  data$tracer <- log2(data$periphery_tracer / data$interior_tracer)
  data$antibody <- log2(data$periphery_antibody / data$interior_antibody)

  data
}

PlotEnrichment <- function(data) {
  # Plot the enrichment ratios for every experiment group
  data <- data[, c("experiment", "tracer", "antibody")]
  
  data <- melt(data, id.vars = "experiment")
  
  plt <- ggplot(data, aes(x = experiment, col = variable, y = value,
                          group = interaction(experiment, variable))) +
    geom_hline(yintercept = 0, col = "black", linetype = "dashed") +
    geom_quasirandom(dodge.width = 0.75, size = 1) +
    geom_boxplot(outlier.shape = NA, fill = NA, col = "black") +
    ggtitle("peripheral-interior enrichment") +
    xlab("") +
    ylab("enrichment (log2)") +
    scale_colour_brewer(palette = "Set1") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))

  plt
}

PlotRawScores <- function(data, channel = "tracer") {
  # Plot the enrichment ratios for every experiment group
  if (channel == "tracer") {
    data <- data[, c("experiment", 
                   "periphery_tracer", "interior_tracer")]
  } else if (channel == "antibody") {
    data <- data[, c("experiment", 
                   "periphery_antibody", "interior_antibody")]
  } else {
    data <- data[, c("experiment", 
                   "periphery_tracer", "interior_tracer",
                   "periphery_antibody", "interior_antibody")]
  }
  
  
  data <- melt(data, id.vars = "experiment")
  
  plt <- ggplot(data, aes(x = experiment, col = variable, y = value,
                          group = interaction(experiment, variable))) +
    geom_quasirandom(dodge.width = 0.75, size = 1) +
    geom_boxplot(outlier.shape = NA, fill = NA, col = "black") +
    ggtitle("raw scores") +
    xlab("") +
    ylab("mean intensity") +
    scale_colour_brewer(palette = "Set1",
                        labels = c("periphery", "interior")) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))

  plt
}

```

### Load data

Read in the data, given a list of directories with "analysis" results from the 
imageJ script that creates 4 .csv files for every slide.

```{r load data}

# Input directory for the files
directories <- c("./data/pADamID_Hap1_plus6_mouseAntibodies_analysis/",
                 
                 "./data/pADamID_MethylTracer_analysis/",
                 
                 "./data/ts181030_pADamID_M6Tracer_analysis/",
                 
                 "./data/ts181030_pADamID_stability_analysis/")



data <- c()

for (directory in directories) {
  # Get the slide names
  slides <- ListSlides(directory)

  # Extract experiment names
  experiments <- GetExperiment(slides)
  
  # Read in all the .csv files and form one data frame
  for (i in 1:length(slides)) {
    data <- rbind(data,
                  ReadSlide(slide_name = file.path(directory, slides[i]),
                            experiment = experiments[i]))
  }
}

```

### Calculate enrichment

Calculate the enrichment ratios for every cell.

```{r calculate enrichment}

# Calculate enrichment
data <- CalculateLog2Ratios(data)

```

### Plot enrichment

Finally, create a plot with enrichment scores. Also, write the mean scores for 
each experiment to a file.

```{r plot enrichment, fig.width = 10, fig.height = 5}

plt <- PlotEnrichment(data)
plot(plt)

plt <- PlotRawScores(data)
plot(plt)

```

```{r plot selected enrichments, fig.width = 4.5, fig.height = 3.5}

# Also, create a plot with selected features
data.filtered <- as_tibble(data) %>%
  mutate(experiment = as.character(experiment)) %>%
  mutate(experiment = case_when(experiment == "sample_2" ~ "Dam",
                                experiment == "sample_5-2" ~ "Lamin B1",
                                experiment == "sample_12" ~ "H3K27me3",
                                experiment == "sample_14" ~ "H3K9me3",
                                experiment == "LMNB1_rabbit" ~ "Lamin B1",
                                experiment == "LMNB2" ~ "Lamin B2",
                                experiment == "pA-Dam" ~ "pADam",
                                experiment == "L1-100" ~ "Lamin B1",
                                experiment == "L1-500" ~ "Lamin B1",
                                experiment == "L2-100" ~ "Lamin B2",
                                experiment == "L2-500" ~ "Lamin B2",
                                T ~ experiment)) %>%
  filter(experiment %in% c("Dam", "pADam", "Lamin B1", "Lamin B2", "H3K27me3",
                           "H3K9me3")) %>%
  mutate(experiment = factor(experiment, levels = c("Dam", "pADam", "Lamin B1", 
                                                    "Lamin B2", "H3K27me3", 
                                                    "H3K9me3"))) %>%
  as.data.frame()

plt <- PlotEnrichment(data.filtered)
plot(plt)

plt <- PlotRawScores(data.filtered)
plot(plt)

```

```{r write median}

# Calculate median
median.df <- as.data.frame(tapply(data$tracer, data$experiment, median))
names(median.df) <- "enrichment"
median.df$samples <- rownames(median.df)
write.table(median.df, output_file,
            row.names = F, col.names = T, sep = "\t", quote = F)

```

### SessionInfo

```{r sessionInfo}

sessionInfo()

```


