---
title: "Differential LADs and Hi-C subcompartments & CTCF peaks"
author: "Tom van Schaik"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: 
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---


### Introduction

Reviewer #3 asked how the differential LADs compare with Hi-C subcompartments.
I don't think this is necessary to include, but let's do it anyway.

For this, I need differential LADs and Hi-C subcompartments. I will use the
subcompartments as imputed by SNIPER (Kyle, Ma lab).

ts200326: update, also do this for CTCF peaks.


### Method

Simply overlap of the genome.


### Set-up

Load the libraries and set the parameters.

```{r set-up, message = F}

# Load dependencies
library(tidyverse)
library(ggplot2)
library(GenomicRanges)
library(rtracklayer)
library(RColorBrewer)
library(limma)

# # Prepare output 
output.dir <- "ts200318_hic_subcompartments"
dir.create(output.dir, showWarnings = FALSE)

```

```{r knits setup}
library(knitr)
opts_chunk$set(fig.width = 10, fig.height = 4, dev=c('png', 'pdf'), 
               fig.path = file.path(output.dir, "figures/")) 
pdf.options(useDingbats = FALSE)
```

```{r functions}

CalculateOverlap <- function(LADs, compartments, cell) {
  
  # Convert to factor
  compartments$name <- factor(compartments$name)
  
  # Intersect
  compartments.down <- compartments[compartments %over% LADs[LADs$diff == -1]]
  table.down <- table(compartments.down$name)
  
  compartments.stable <- compartments[compartments %over% LADs[LADs$diff == 0]]
  table.stable <- table(compartments.stable$name)
  
  compartments.up <- compartments[compartments %over% LADs[LADs$diff == 1]]
  table.up <- table(compartments.up$name)
  
  # Combine into tibble
  tib <- tibble(subcompartment = factor(c("A1", "A2", "B1", "B2", "B3")),
                down = as.vector(table.down),
                stable = as.vector(table.stable),
                up = as.vector(table.up)) %>%
    gather(key, value, -subcompartment) %>%
    mutate(key = factor(key, levels = c("down", "stable", "up"))) %>% 
    group_by(key) %>% 
    mutate(perc = value / sum(value)) %>%
    ungroup() %>%
    add_column(cell = cell)
  
  tib
  
}

PlotFractions <- function(ovl, main = "", combined = F) {
  
  plt.base <- ovl %>%
    ggplot(aes(x = key, fill = subcompartment)) +
    xlab("") +
    ggtitle(main) +
    scale_fill_manual(values = c("darkgreen", "lightgreen",
                                 "red", "yellow", "darkgrey")) +
    theme_bw() +
    theme(aspect.ratio = 1)
  
  if (combined) {
    plt.base <- plt.base +
      facet_grid(. ~ cell) +
      theme(axis.text.x = element_text(angle = 90, hjust = 1),
            aspect.ratio = 3)
  }
  
  # Counts
  plt <- plt.base +
    geom_bar(aes(y = value), stat = "identity") +
    ylab("# 100kb bins")
  
  plot(plt)
  
  # Percentages
  plt <- plt.base +
    geom_bar(aes(y = perc), stat = "identity") +
    ylab("Fraction")
  
  plot(plt)
    
}

CalculateOverlapInverted <- function(LADs, compartments) {
  
  # Convert to factor
  compartments$name <- factor(compartments$name)
  
  # Intersect
  ovl <- findOverlaps(compartments, LADs, type = "any", select = "arbitrary")
  
  compartments$lad <- LADs$diff[ovl]
  
  # Change into words
  compartments$lad[is.na(compartments$lad)] <- "non-LAD"
  compartments$lad[compartments$lad == -1] <- "down"
  compartments$lad[compartments$lad == 0] <- "stable"
  compartments$lad[compartments$lad == 1] <- "up"
  
  compartments$lad <- factor(compartments$lad, levels = c("down", "stable",
                                                          "up", "non-LAD"))
  
  # Table
  tib <- as_tibble(mcols(compartments)) %>%
    dplyr::select(name, lad) %>%
    group_by(name, lad) %>%
    dplyr::summarise(value = n()) %>%
    group_by(name) %>%
    mutate(perc = value / sum(value)) %>%
    ungroup()
  
  tib
  
}

PlotFractionsInverted <- function(ovl, main = "") {
  
  plt.base <- ovl %>%
    ggplot(aes(x = name, fill = lad)) +
    xlab("") +
    ggtitle(main) +
    scale_fill_manual(values = c(brewer.pal(9, "Set1")[c(1, 9, 3)],
                                 "snow2")) +
    theme_bw() +
    theme(aspect.ratio = 1)
  
  # Counts
  plt <- plt.base +
    geom_bar(aes(y = value), stat = "identity") +
    ylab("# 100kb bins")
  
  plot(plt)
  
  # Percentages
  plt <- plt.base +
    geom_bar(aes(y = perc), stat = "identity") +
    ylab("Fraction")
  
  plot(plt)
    
}

```


### 1. Read data

First the subcompartments. Downloaded from the article.

```{r hic subcompartments}

# Read SNIPER output
compartment.hap1 <- import("ts200318_hic_subcompartments/HAP1_track_hg38.bed.gz")
compartment.k562 <- import("ts200318_hic_subcompartments/K562_track_hg38.bed.gz")

```

Next, the differential LADs.

```{r read LADs}

# Read differential LADs
lads.hap1 <- import("../ts190301_pADamID_CellCycle/ts190708_differential_analysis_BinsandLADs_Hap1/LADs.bed")
lads.results.hap1 <- readRDS("../ts190301_pADamID_CellCycle/ts190708_differential_analysis_BinsandLADs_Hap1/LADs_results.rds")

lads.hap1$diff <- lads.results.hap1[, 1]

lads.k562 <- import("../ts190301_pADamID_CellCycle/ts190708_differential_analysis_BinsandLADs_K562/LADs.bed")
lads.results.k562 <- readRDS("../ts190301_pADamID_CellCycle/ts190708_differential_analysis_BinsandLADs_K562/LADs_results.rds")

lads.k562$diff <- lads.results.k562[, 1]

lads.hct116 <- import("../ts190301_pADamID_CellCycle/ts190708_differential_analysis_BinsandLADs_HCT116/LADs.bed")
lads.results.hct116 <- readRDS("../ts190301_pADamID_CellCycle/ts190708_differential_analysis_BinsandLADs_HCT116/LADs_results.rds")

lads.hct116$diff <- lads.results.hct116[, 1]

```


### 2. Overlap between LADs and subcompartments

With the LADs and compartments, I can compute the overlap.

```{r calculate overlap}

ovl.hap1 <- CalculateOverlap(lads.hap1, compartment.hap1, cell = "HAP-1")
ovl.k562 <- CalculateOverlap(lads.k562, compartment.k562, cell = "K562")

```


### 3. Plot

Plot.

```{r plot, fig.width = 5, fig.height = 3.5}

# Various plots
PlotFractions(ovl.hap1, main = "HAP-1")
PlotFractions(ovl.k562, main = "K562")

```


```{r combined plot, fig.width = 8, fig.height = 3.5}

# Combined plot
PlotFractions(bind_rows(ovl.hap1, ovl.k562), combined = T)

```


### 4. Repeat #2 and #3, but inverted

```{r overlap 2}

ovl.inverse.hap1 <- CalculateOverlapInverted(lads.hap1, compartment.hap1)
ovl.inverse.k562 <- CalculateOverlapInverted(lads.k562, compartment.k562)

```

```{r plot 2, fig.width = 5, fig.height = 3.5}

# Various plots
PlotFractionsInverted(ovl.inverse.hap1, main = "Hap1")
PlotFractionsInverted(ovl.inverse.k562, main = "K562")

```


### 5. Repeat, for CTCF peaks

### 5.1. Read data

Read CTCF peaks.

```{r ctcf peaks}

# Read SNIPER output
peaks.hap1 <- import("ts200318_hic_subcompartments/ts200326_Hap1_ChIP_CTCF_hg38_GSM2493878_peaks.narrowPeak.gz")
peaks.k562 <- import("ts200318_hic_subcompartments/ts200326_K562_ChIP_CTCF_ENCFF519CXF.narrowPeak.gz")
peaks.hct116 <- import("ts200318_hic_subcompartments/ts200326_HCT116_ChIP_CTCF_ENCFF518MQA.narrowPeak.gz")

```


### 2. Overlap between LADs and subcompartments

With the LADs and compartments, I can compute the overlap.

```{r calculate overlap peaks}

# Chrom sizes
chrom.sizes <- read.table("~/mydata/data/genomes/GRCh38/hg38.chrom.sizes", sep = "\t")
row.names(chrom.sizes) <- chrom.sizes[, 1]

chrom.sizes <- chrom.sizes[c(paste0("chr", 1:22), "chrX"), ]

CalculateOverlapPeaks <- function(LADs, peaks, cell) {
  
  # Add chrom sizes
  seqlengths(LADs) <- chrom.sizes[seqlevels(LADs), 2]
  
  # Get LAD sizes
  LAD_class.widths <- c(sum(width(LADs[LADs$diff == -1])),
                        sum(width(LADs[LADs$diff == 0])),
                        sum(width(LADs[LADs$diff == 1])))
  LAD_class.widths[4] <- sum(chrom.sizes[, 2]) - sum(LAD_class.widths)
  
  # Intersect peaks with LADs
  ovl <- findOverlaps(peaks, LADs, type = "any", select = "arbitrary")
  peaks.lad_status <- LADs$diff[ovl]
  
  # Combine into tibble
  tib <- tibble(class = c("down", "stable", "up", "iLAD"),
                peaks = c(sum(peaks.lad_status == -1, na.rm = T), 
                          sum(peaks.lad_status == 0, na.rm = T),
                          sum(peaks.lad_status == 1, na.rm = T),
                          sum(is.na(peaks.lad_status))),
                widths = LAD_class.widths / 1e6) %>%
    mutate(class = factor(class, levels = c("down", "stable", "up", "iLAD"))) %>% 
    mutate(density = peaks / widths) %>%
    add_column(cell = cell)
  
  tib
  
}

ovl.hap1 <- CalculateOverlapPeaks(lads.hap1, peaks.hap1, cell = "HAP-1")
ovl.k562 <- CalculateOverlapPeaks(lads.k562, peaks.k562, cell = "K562")
ovl.hct116 <- CalculateOverlapPeaks(lads.hap1, peaks.hct116, cell = "HCT116")

```


### 3. Plot

Plot.

```{r plot peaks, fig.width = 5, fig.height = 3.5}

# Various plots
PlotFractionsPeaks <- function(ovl, main = "", combined = F) {
  
  plt.base <- ovl %>%
    ggplot(aes(x = class, fill = class)) +
    xlab("") +
    ggtitle(main) +
    scale_fill_brewer(palette = "Set2") +
    theme_bw() +
    theme(aspect.ratio = 1)
  
  if (combined) {
    plt.base <- plt.base +
      facet_grid(. ~ cell) +
      theme(axis.text.x = element_text(angle = 90, hjust = 1),
            aspect.ratio = 2)
  }
  
  # Density
  plt <- plt.base +
    geom_bar(aes(y = density), stat = "identity") +
    ylab("#CTCF peaks / Mb")
  
  plot(plt)
  
}

PlotFractionsPeaks(ovl.hap1, main = "HAP-1")
PlotFractionsPeaks(ovl.k562, main = "K562")
PlotFractionsPeaks(ovl.hct116, main = "HCT116")

```


```{r combined plot peaks, fig.width = 8, fig.height = 3.5}

# Combined plot
PlotFractionsPeaks(bind_rows(ovl.hap1, ovl.k562, ovl.hct116), combined = T)

```


### Conclusion

There is a small increase in the B1 subcompartment in decreasing LADs. But I
don't think that this significant and worth including.


### SessionInfo

```{r sessioninfo}

sessionInfo()

```
